Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-26 prop_letter.tex
revision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Born rule and Gleason's Theorem}

@@

\noindent For general QIVPMs mapping to finitely precise intervals, we
also want to seek a simple rule to construct them, and to understand
which states are consistent with them. However, if measurement
intervals are too imprecise, the state consistent with the results of
measurements on some projectors might contradict the state induced by
some other projectors. Indeed we can formally prove that, in general,
there may be \emph{no} state consistent with some arbitrary QIVPM.

@@

To prove the theorem, we first need to construct a QIVPM, and then
verify that there are no states that are consistent with it on all
projectors. Given an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $, let
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined in
Table~\ref{tab:non-Born-QIVPM} can be verified as a
QIVPM~\citep{HOSTunpublished}.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on a Hilbert space of dimension~$d\ge3$. The projectors mapping
to the same interval are listed in the column labeled by $P$ and
separated by commas.}

@@

To that end, we construct a family of QIVPMs parameterized by a
functions $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$ that
formalize how the real-valued probabilities are mapped to
intervals. These functions, are called \emph{ultramodular}, when they
satisfy certain properties defined next.

\begin{definition}[Ultramodular Functions]\label{def:THOS}Given
a collection of intervals $\mathscr{I}$ including $\imposs$ and
$\necess$, a function $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$
is called a ultramodular function if:
\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \ultramodular(0)=\imposs\,,\\
 &  & \ultramodular(1)=\necess\,,\\
 &  & \ultramodular\left(1-x\right)=\necess-\ultramodular\left(x\right)\,,
\end{eqnarray}
\end{subequations}and for any three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that
$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, we have:
\begin{equation}
\ultramodular\left(x_{0}+x_{1}+x_{2}\right)+\ultramodular\left(x_{2}\right)\subseteq\ultramodular\left(x_{0}+x_{2}\right)+\ultramodular\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent The first three constraints,
Eqs.~(\ref{eq:iota-constraints}), are the direct counterpart of the
corresponding QIVPM constraints, Eqs.~(\ref{eq:QIVPM-constraints});
the last condition, Eq.~(\ref{eq:iota-Inclusion}), is the direct
counterpart of the convexity conditions,
Eq.~(\ref{eq:classicalconvex}) and
(\ref{eq:QuantumInterval-valuedProbability-Inclusion}). Therefore,
these conditions guarantee that, for any conventional quantum
probability measure $\mu$, the composition $\ultramodular \circ \mu$,
defines a valid QIVPM. Conversely, it is the case that if
$f \circ \mu$ is a QIVPM for some function $f$ then $f$ is an
ultramodular function. Formally, we have the following result.

@@

\begin{thm}[Equivalence of Ultramodular Functions and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}A function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$
is ultramodular.
\item \label{enu:iota-mu-CIVPM}The composite
  function~$\ultramodular\circ\mu:\events_{\mathrm{C}}\rightarrow\mathscr{I}$
  is a classical IVPM for any classical probability
  measure~$\mu:\events_{\mathrm{C}}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composite
  function~$\ultramodular\circ\mu:\events\rightarrow\mathscr{I}$ is a
  QIVPM for any quantum probability
  measure~$\mu:\events\rightarrow\left[0,1\right]$ on any Hilbert
  space~$\Hilb$.
\end{enumerate}
\end{thm}

The equivalence between Statement~\ref{enu:iota-subject-to} and
\ref{enu:iota-mu-CIVPM} is a well-known classical
result~\citep{Choquet1954,NgMoYeh1997,Marinacci_2004}.  For the
quantum case, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM} as we have outlined above. The rest of the
proof is to show that if $\ultramodular$ is not ultramodular, then
$\ultramodular\circ\mu$ might not be a QIVPM. Suppose there are three
particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but they don't
satisfy Eq.~(\ref{eq:iota-Inclusion}). Consider the state:
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\,.
\end{aligned}
\end{equation}
The induced map using the Born rule and approximated by
$\ultramodular$, $\ultramodular\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$
fails to satisfy
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}) when
$P_{0}=\proj{0}+\proj{2}$ and $P_{1}=\proj{1}+\proj{2}$. In other
words, this induced map fails to be a QIVPM.

In contrast with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM},
there is always a state consistent with any composite QIVPM
$\ultramodular\circ\mu_{\rho}^{\mathrm{B}}$.  Moreover, the collection
of states consistent with $\ultramodular\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to~$\rho$ as the intervals get narrower and
narrower. Eventually, the collection of states collapses to the
singleton set~$\left\{ \rho\right\} $ as the intervals become the
infinitely precise intervals~$\mathscr{I}_{\infty}$. To precisely
state this result, we introduce the norm of pure states and
operators. The norm of a pure state~$\ket{\psi}$ is defined as usual
$\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.  Similarly,
given any Hermitian operator~$A$, we consider the usual operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert
  =1}\left\Vert A\ket{\psi}\right\Vert $, which is also known as the
$2$-norm or the spectral
norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
Then, a version of Gleason's theorem can be stated as follow.

\begin{thm}[Finite-Precision Extension of Gleason's
  theorem]\label{thm:Finite-precision-Gleason}Consider
  an ultramodular function
  $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$ and let $m$
  be the maximum length of intervals in its image. Given any
  state~$\rho$, if a state
  $\rho'\in\coreBorn\left(\ultramodular\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
  then the norm of their difference is bounded by $m$, i.e.,
  $\left\Vert \rho-\rho'\right\Vert \le m$.\end{thm}

Given a state~$\rho'$ consistent with $\ultramodular\circ\mu_{\rho}^{\mathrm{B}}$,
we have $\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\ultramodular\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$
for any one-dimensional projector $\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $m$, it is also the
upper bound of the difference:
\[
\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le m\,.
\]
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le m$.

The previous theorem generalizes Gleason's theorem in the sense that
it accounts for a larger class of probability measures which includes
the conventional one as a limit. The theorem is however ``weak'' in
the sense that it only applies to the particular class of QIVPMs
constructed by composing an ultramodular function with a conventional
quantum probability measure.

\amr{***WAVEFRONT***}

Although a ultramodular function has many desirable properties, the
set of intervals it mapping to either is trivial or contains infinite
intervals.

@@

Since $\ultramodular$ maps to intervals, we can decompose it into
two functions: its left-end and right-end $\left[\ultramodularL\left(x\right),\ultramodularR\left(x\right)\right]=\ultramodular\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function $\ultramodularL:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992},
i.e., 
\begin{equation}
\begin{aligned} & \ultramodularL\left(x_{0}+x_{1}+x_{2}\right)+\ultramodularL\left(x_{2}\right)\\
\ge{} & \ultramodularL\left(x_{0}+x_{2}\right)+\ultramodularL\left(x_{1}+x_{2}\right)\,.
\end{aligned}
\end{equation}
Then, for any $x_{0}'$ and $x_{1}'\in\left[0,1\right]$ with $x_{0}'\ge x_{1}'$,
we can substitute $x_{0}$ and $x_{1}$ by $\left(x_{0}'-x_{1}'\right)/2$
, substitute $x_{2}$ by $x_{1}'$, and get the following inequality
\begin{eqnarray}
\frac{\ultramodularL\left(x_{0}'\right)+\ultramodularL\left(x_{1}'\right)}{2} & \ge & \ultramodularL\left(\frac{x_{0}'+x_{1}'}{2}\right)\,,
\end{eqnarray}
i.e., $\ultramodularL$ is midpoint-convex or Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992}.
Together with the fact that $\ultramodularL$ maps to a bounded interval
$\left[0,1\right]$, the left-end function~$\ultramodularL$ must
be continuous on the unit open interval $\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.
Therefore, either $\ultramodular$ maps every number in $\left(0,1\right)$
to the same interval, or the number of intervals that $\ultramodular$
maps to must be uncountable.

Even if $\ultramodular$ has to be ultramodular to guarantee $\ultramodular\circ\mu$
a QIVPM, the composited function $\ultramodular\circ\mu$ may still
be a QIVPM when $\ultramodular$ is not ultramodular. In this case,
the composited function $\ultramodular\circ\mu$ could map to non-trivial
finite number of intervals. For example, we can weaken the convex
condition, Eq.~(\ref{eq:iota-Inclusion}), to the following additive
condition: for any numbers $x_{0}$ and $x_{1}\in\left[0,1\right]$,
their sum $x_{0}+x_{1}\in\left[0,1\right]$ implies
\begin{equation}
\ultramodular\left(x_{0}+x_{1}\right)\subseteq\ultramodular\left(x_{0}\right)+\ultramodular\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, given a three-dimensional Hilbert space, when $\ultramodular$
satisfies Eqs.~(\ref{eq:iota-constraints}) and Eq.~(\ref{eq:iota-add}),
$\ultramodular\circ\mu$ still must be a QIVPM~\citep{HOSTunpublished}.
More concretely, consider an orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$,
two states $\rho_{0}=\frac{1}{3}\proj{0}+\frac{1}{3}\proj{1}+\frac{1}{3}\proj{2}$
and $\rho_{1}=\frac{7}{24}\proj{0}+\frac{1}{3}\proj{1}+\frac{3}{8}\proj{2}$,
and a set of intervals
\begin{equation}
\begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
 & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{3}{4},1\right],\necess\right\} \,.
\end{aligned}
\end{equation}
Since the interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
and Eq.~(\ref{eq:iota-add})~\citep{HOSTunpublished}, $\ultramodular_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
and $\ultramodular_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ are both
QIVPMs.
\begin{table}
\caption{\label{tab:I_2}A interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Born rule and Gleason's Theorem}

@@

\noindent For general QIVPMs mapping to imprecise intervals, we also
want to seek a simple rule to construct them, and to understand which
states are consistent with them. However, if measurement intervals are
too imprecise, the state consistent with the results of measurements
on some projectors might contradict the state induced by some other
projectors. Indeed we can formally prove that, in general, there may
be \emph{no} state consistent with some arbitrary QIVPM.

@@

To prove the theorem, we just need to construct a QIVPM, and verify
that there are no states that are consistent with it on all
projectors. Given an orthonormal basis
$\left\{ \ket{0},\ket{1},\ket{2},\ldots,\ket{d-1}\right\} $, let
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{equation}
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.\label{eq:3-value-intervals}
\end{equation}
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined in
Table~\ref{tab:non-Born-QIVPM} can be verified as a
QIVPM~\citep{HOSTunpublished}.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on a Hilbert space of dimension~$d\ge3$. The projectors mapping
to the same interval are listed in the column labeled by $P$ and
separated by commas.}

@@

To that end, we construct a family of QIVPMs parameterized by
clustering functions $f:\left[0,1\right]\rightarrow\mathscr{I}$ that
formalize how the real-valued probabilities are mapped to
intervals. We also introduce a notion of norm that quantifies the
distance between (pure or mixed) states.  The norm of a pure
state~$\ket{\psi}$ is defined as usual
$\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.  Similarly,
given any Hermitian operator~$A$, we consider the usual operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert
  =1}\left\Vert A\ket{\psi}\right\Vert $, which is also known as the
$2$-norm or the spectral
norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
Then, a version of Gleason's theorem can be stated as follows.

@@

\begin{thm}[Finite-Precision Extension of Gleason's
  theorem]\label{thm:Finite-precision-Gleason}Let $f$ be a clustering
  function and let $f \circ \mu_{\rho}^{\mathrm{B}}$ be a QIVPM, where
  $\mu_{\rho}^{\mathrm{B}}$ is the probability measure induced by the
  Born rule for a given state $\rho$. Let $\alpha$ be the maximum
  length of intervals in its image. If a state $\rho'$ is consistent
  with $f\circ\mu_{\rho}^{\mathrm{B}}$ on all events, i.e.,
  $\rho'\in\coreBorn\left(f\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
  then the norm of their difference is bounded by $\alpha$, i.e.,
  $\left\Vert \rho-\rho'\right\Vert \le \alpha$.\end{thm}

The proof proceeds as follows. Given a state~$\rho'$ consistent with
$f\circ\mu_{\rho}^{\mathrm{B}}$, we have
$\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in
f\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$ for any
one-dimensional projector $\proj{\psi}$. Since the maximum length of
the intervals in $\mathscr{I}$ is $\alpha$, it is also the upper bound
of the difference:
\[
\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le \alpha\,.
\]
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le \alpha$.

The previous theorem generalizes Gleason's theorem in the sense that
it accounts for a larger class of probability measures which includes
the conventional one as a limit. The theorem is however ``special'' in
the sense that it only applies to the particular class of QIVPMs
constructed by composing a clustering function with a conventional
quantum probability measure. QIVPMs constructed in this manner have
some peculiar properties which we examine next.

A clustering function is called \emph{ultramodular} if it satisfies
certain properties.

\begin{definition}[Ultramodular Functions]\label{def:THOS}Given
  a collection of intervals $\mathscr{I}$ including $\imposs$ and
  $\necess$, a clustering function
  $\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$ is called
  ultramodular if:
\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \ultramodular(0)=\imposs\,,\\
 &  & \ultramodular(1)=\necess\,,\\
 &  & \ultramodular\left(1-x\right)=\necess-\ultramodular\left(x\right)\,,
\end{eqnarray}
\end{subequations}and for any three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that
$y=x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, we have:
\begin{equation}
\ultramodular\left(y\right)+\ultramodular\left(x_{2}\right)\subseteq\ultramodular\left(x_{0}+x_{2}\right)+\ultramodular\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

\noindent The first three constraints,
Eqs.~(\ref{eq:iota-constraints}), are the direct counterpart of the
corresponding QIVPM constraints, Eqs.~(\ref{eq:QIVPM-constraints});
the last condition, Eq.~(\ref{eq:iota-Inclusion}), is the direct
counterpart of the convexity conditions,
Eq.~(\ref{eq:classicalconvex}) and
(\ref{eq:QuantumInterval-valuedProbability-Inclusion}). Therefore,
these conditions guarantee that, for any conventional quantum
probability measure $\mu$, the composition $\ultramodular \circ \mu$,
defines a valid QIVPM. Conversely, if for every quantum probability
measure $\mu$, it is the case that $f \circ \mu$ is a QIVPM, then $f$
is an ultramodular function. Formally, we have the following result.

\begin{thm}[Equivalence of Ultramodular Functions and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}A function~$\ultramodular:\left[0,1\right]\rightarrow\mathscr{I}$
is ultramodular.
\item \label{enu:iota-mu-CIVPM}The composite
  function~$\ultramodular\circ\mu:\events_{\mathrm{C}}\rightarrow\mathscr{I}$
  is a classical IVPM for all classical probability
  measure~$\mu:\events_{\mathrm{C}}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composite
  function~$\ultramodular\circ\mu:\events\rightarrow\mathscr{I}$ is a
  QIVPM for all quantum probability
  measure~$\mu:\events\rightarrow\left[0,1\right]$.
\end{enumerate}
\end{thm}

The equivalence between Statement~\ref{enu:iota-subject-to} and
\ref{enu:iota-mu-CIVPM} is a well-known classical
result~\citep{Choquet1954,NgMoYeh1997,Marinacci_2004}.  For the
quantum case, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM} as we have outlined above. The rest of the
proof is to show that if $\ultramodular$ is not ultramodular, then for
some quantum probability measure $\mu$, the composite
$\ultramodular\circ\mu$ might not be a QIVPM. Suppose there are three
particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but they don't
satisfy Eq.~(\ref{eq:iota-Inclusion}). Consider the state:
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\,.
\end{aligned}
\end{equation}
The induced map using the Born rule and clustered by $\ultramodular$,
$\ultramodular\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ fails to satisfy
Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}) when
$P_{0}=\proj{0}+\proj{2}$ and $P_{1}=\proj{1}+\proj{2}$. In other
words, this induced map fails to be a QIVPM.

% In contrast with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM},
% there is always a state consistent with any composite QIVPM
% $\ultramodular\circ\mu_{\rho}^{\mathrm{B}}$.  Moreover, the collection
% of states consistent with $\ultramodular\circ\mu_{\rho}^{\mathrm{B}}$
% gets closer and closer to~$\rho$ as the intervals get narrower and
% narrower. Eventually, the collection of states collapses to the
% singleton set~$\left\{ \rho\right\} $ as the intervals become the
% infinitely precise intervals~$\mathscr{I}_{\infty}$. To precisely
% state this result, we introduce the norm of pure states and
% operators. The norm of a pure state~$\ket{\psi}$ is defined as usual
% $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.  Similarly,
% given any Hermitian operator~$A$, we consider the usual operator norm
% $\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert
%   =1}\left\Vert A\ket{\psi}\right\Vert $, which is also known as the
% $2$-norm or the spectral
% norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
% Then, a version of Gleason's theorem can be stated as follow.

In other words, essential properties of QIVPMs constructed using
clustering functions can be gleaned from the properties of
ultramodular functions. The following is a most interesting property
in our setting.

@@

Since $\ultramodular$ maps to intervals, we can decompose it into
two functions: its left-end and right-end $\left[\ultramodularL\left(x\right),\ultramodularR\left(x\right)\right]=\ultramodular\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function $\ultramodularL:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992},
i.e., 
\begin{equation}
\begin{aligned} & \ultramodularL\left(x_{0}+x_{1}+x_{2}\right)+\ultramodularL\left(x_{2}\right)\\
\ge{} & \ultramodularL\left(x_{0}+x_{2}\right)+\ultramodularL\left(x_{1}+x_{2}\right)\,.
\end{aligned}
\end{equation}
Then, for any $x_{0}'$ and $x_{1}'\in\left[0,1\right]$ with $x_{0}'\ge x_{1}'$,
we can substitute $x_{0}$ and $x_{1}$ by $\left(x_{0}'-x_{1}'\right)/2$
, substitute $x_{2}$ by $x_{1}'$, and get the following inequality
\begin{eqnarray}
\frac{\ultramodularL\left(x_{0}'\right)+\ultramodularL\left(x_{1}'\right)}{2} & \ge & \ultramodularL\left(\frac{x_{0}'+x_{1}'}{2}\right)\,,
\end{eqnarray}
i.e., $\ultramodularL$ is midpoint-convex or
Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992}.
Together with the fact that $\ultramodularL$ maps to a bounded
interval $\left[0,1\right]$, the left-end function~$\ultramodularL$
must be continuous on the unit open interval
$\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.  Therefore,
either $\ultramodular$ maps every number in $\left[0,1\right]$ to the
same interval, or the number of intervals that $\ultramodular$ maps to
must be uncountable.

To summarize, a conventional quantum probability measure, has an
uncountable range $[0,1]$. A QIVPM constructed by clustering such a
conventional quantum probabily measure, must also have an uncountable
range of intervals. Of course, any particular QIVPM, or any particular
experiment, will use a fixed collection of intervals appropriate for
the resources and precision of the particular experiment.

% The essential reason for getting an uncountable number of intervals is
% the convexity condition we imposed on QIVPMs. This condition is widely
% used in the classical setting but it is possible that it should be
% modified or weakened in the quantum case. If we take this approach and 
% % Even if $\ultramodular$ has to be ultramodular to guarantee
% % $\ultramodular\circ\mu$ a QIVPM, the composited function
% % $\ultramodular\circ\mu$ may still be a QIVPM when $\ultramodular$ is
% % not ultramodular. In this case, the composited function
% % $\ultramodular\circ\mu$ could map to non-trivial finite number of
% % intervals. For example, we can 
% weaken the convexity condition, Eq.~(\ref{eq:iota-Inclusion}), we can
% recover QIVPMs that only require countable and even finite collections
% of intervals. Formally, we weaken the convexity condition to the
% following additive condition: for any numbers $x_{0}$ and
% $x_{1}\in\left[0,1\right]$, their sum $x_{0}+x_{1}\in\left[0,1\right]$
% implies
% \begin{equation}
% \ultramodular\left(x_{0}+x_{1}\right)\subseteq\ultramodular\left(x_{0}\right)+\ultramodular\left(x_{1}\right)\,.\label{eq:iota-add}
% \end{equation}
% Then, given a three-dimensional Hilbert space, if $\ultramodular$
% satisfies Eqs.~(\ref{eq:iota-constraints}) and
% Eq.~(\ref{eq:iota-add}), then $\ultramodular\circ\mu$ is a
% QIVPM~\citep{HOSTunpublished}.  For example, consider an orthonormal
% basis $\{\ket{0},\ket{1},\ket{2}\}$, two states
% $\rho_{0}=\frac{1}{3}\proj{0}+\frac{1}{3}\proj{1}+\frac{1}{3}\proj{2}$
% and
% $\rho_{1}=\frac{7}{24}\proj{0}+\frac{1}{3}\proj{1}+\frac{3}{8}\proj{2}$,
% and a set of intervals:
% \begin{equation}
% \begin{aligned}\mathscr{I}_{2}={} & \left\{ \imposs,\left[0,\tfrac{1}{4}\right],\left[0,{\textstyle \tfrac{1}{2}}\right],\left[\tfrac{1}{4},\tfrac{3}{4}\right],\right.\\
%  & \left.\left[\tfrac{1}{2},1\right],\left[\tfrac{3}{4},1\right],\necess\right\} \,.
% \end{aligned}
% \end{equation}
% Since the interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$
% defined in Table~\ref{tab:I_2} satisfies Eqs.~(\ref{eq:iota-constraints})
% and Eq.~(\ref{eq:iota-add})~\citep{HOSTunpublished}, $\ultramodular_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
% and $\ultramodular_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$ are both
% QIVPMs.
% \begin{table}
% \caption{\label{tab:I_2}A interval-valued function~$\ultramodular_{2}:\left[0,1\right]\rightarrow\mathscr{I}_{2}$.}

% \hfill{}%
% \begin{minipage}[t]{2.5cm}%
% \begin{center}
% \begin{tabular}{cc}
% \toprule 
% \addlinespace
% $x$ & $\ultramodular_{2}\left(x\right)$\tabularnewline\addlinespace
% \midrule
% \midrule 
% \addlinespace
% $x=0$ & $\imposs$\tabularnewline\addlinespace
% \midrule 
% \addlinespace
% $0<x\le\tfrac{1}{8}$ & $\left[0,\tfrac{1}{4}\right]$\tabularnewline\addlinespace
% \midrule 
% \addlinespace
% $\tfrac{1}{8}<x<\tfrac{3}{8}$ & $\left[0,{\textstyle \tfrac{1}{2}}\right]$\tabularnewline\addlinespace
% \midrule 
% \addlinespace
% $\tfrac{3}{8}\le x\le\tfrac{5}{8}$ & $\left[\tfrac{1}{4},\tfrac{3}{4}\right]$\tabularnewline\addlinespace
% \bottomrule
% \end{tabular}
% \par\end{center}%
% \end{minipage}\hfill{}%
% \begin{minipage}[t]{2.5cm}%
% \begin{center}
% \begin{tabular}{cc}
% \toprule 
% \addlinespace
% $x$ & $\ultramodular_{2}\left(x\right)$\tabularnewline\addlinespace
% \midrule
% \midrule 
% \addlinespace
% $\tfrac{5}{8}<x<\tfrac{7}{8}$ & $\left[\tfrac{1}{2},1\right]$\tabularnewline\addlinespace
% \midrule 
% \addlinespace
% $\tfrac{7}{8}\le x<1$ & $\left[\tfrac{3}{4},1\right]$\tabularnewline\addlinespace
% \midrule 
% \addlinespace
% $x=1$ & $\necess$\tabularnewline\addlinespace
% \bottomrule
% \end{tabular}
% \par\end{center}%
% \end{minipage}\hfill{}
% \end{table}
% Given any one-dimensional projector $\proj{\psi}$, we have $\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)=\frac{1}{3}$
% so that $\ultramodular_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)\right)=\left[0,{\textstyle \tfrac{1}{2}}\right]$.
% On the other hand, 
% \begin{eqnarray*}
% \mu_{\rho_{1}}^{\mathrm{B}}\left(\proj{\psi}\right) & = & \tfrac{7}{24}\left|\ip{\psi}{0}\right|^{2}+\tfrac{1}{3}\left|\ip{\psi}{1}\right|^{2}+\tfrac{3}{8}\left|\ip{\psi}{2}\right|^{2}
% \end{eqnarray*}
% is bounded between $\frac{7}{24}$ and $\frac{3}{8}$. Therefore,
% $\mu_{\rho_{1}}^{\mathrm{B}}\left(\proj{\psi}\right)\in\ultramodular_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$.
% According to the complement condition, Eq.~(\ref{eq:complement}), we
% also have
% $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\in\ultramodular_{2}\left(\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)\right)$
% for any two-dimensional projector~$P$. Hence,
% $\rho_{1}\in\coreBorn\left(\ultramodular_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
% and Thm.~\ref{thm:Finite-precision-Gleason} asserts that
% $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is less than or equal to the
% maximum length of intervals $\frac{1}{2}$.  This is true because the
% exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is
% $\frac{1}{24}$.
